Linear Ordered Attribute Grammar scheduling using SAT-solving

### 2014-10-18 paper/paper.tex
kleine nuance
Last spellcheck, fix overfull hbox and remove last comment. This is the version to be submitted!

To find the total order we add edges to the nonterminal dependency graph and rely on the SAT-solver to determine their direction. Instead of adding an edge for all possible pairs of attributes it is sufficient to add an edge for all pairs of inherited and synthesized attributes.
The order described by an assignment to these variables is not (always) total as there might be pairs of attributes, which attributes are either both inherited or both synthesized, without a relative ordering. This is not a problem as their evaluation order can be chosen arbitrarily\Thomas{Dit beschouw ik zelf als een belangrijk resultaat. Het is de uitkomst van het lange proces waarin ik geprobeerd heb de minimale definitie van het probleem van LOAG scheduling te geven. Maar om dit beter te kunnen uitleggen heb ik 'visits' nodig.}.

@@

The above algorithm features a complicated combination of counting networks; one network for each production rule, and one network for each corresponding output of these networks. Still, the procedure finds optimal solutions very quickly in practice, in times that are negligable and not practically measurable compared to the time of generating the initial SAT-problem. The number of iterations for the minimisation loops has never been more than 5 in any of our problems.

The algorithm is guaranteed to find the global optimum. For our largest example, the solution found had a total of 130 visits, which was 29 visits less in total than the previously known optimimum, found using backtracking heuristics.

@@

In the translation into the SAT-problem we have used a novel technique for ruling out cycles in the SAT problem using chordal graphs. The technique was inspired by existing work on chordality for expressing equality logic \cite{bryant-chordal}. Using chordal graphs makes the problems much smaller than directly ruling out cycles, while encoding the same restrictions. We believe that this technique is applicable in other problems using SAT-solvers as well.

>>>>

To find the total order we add edges to the nonterminal dependency graph and rely on the SAT-solver to determine their direction. Instead of adding an edge for all possible pairs of attributes it is sufficient to add an edge for all pairs of inherited and synthesized attributes.
The order described by an assignment to these variables is not (always) total as there might be pairs of attributes, which attributes are either both inherited or both synthesized, without a relative ordering. This is not a problem as the evaluation order for such pairs can be chosen arbitrarily.

@@

The above algorithm features a complicated combination of counting networks; one network for each production rule, and one network for each corresponding output of these networks. Still, the procedure finds optimal solutions very quickly in practice, in times that are negligible and not practically measurable compared to the time of generating the initial SAT-problem. The number of iterations for the minimisation loops has never been more than 5 in any of our problems.

The algorithm is guaranteed to find the global optimum. For our largest example, the solution found had a total of 130 visits, which was 29 visits less in total than the previously known optimum, found using backtracking heuristics.

@@

In the translation into the SAT-problem we have used a novel technique for ruling out cycles in the SAT problem using chordal graphs. Existing work on chordality for expressing equality logic \cite{bryant-chordal} was the inspiration for this technique. Using chordal graphs makes the problems much smaller than directly ruling out cycles, while encoding the same restrictions. We believe that this technique is applicable in other problems using SAT-solvers as well.
